---
title: "Leaflet geoplot"
output:
  html_document:
    df_print: paged
---

In this notebook, we'll use the Leaflet library to plot interactive plots.

## Load libraries

```{r}
library(leaflet)
library(sf)
```

## Load dataset

```{r}
dataset <- read.csv("data/dataset.csv", skip = 4)
dataset[, c("Indicator.Name", "Indicator.Code", "X1960", "X2016", "X")] <- NULL
dataset <- dataset[complete.cases(dataset), ]
dataset["avgPopulationDensity"] <- rowSums(dataset[,3:57])
head(dataset)
```

We also load in the world shapefile and merge the two files to create a common dataset.

```{r}
shapes <- st_read("data/countries_shp/countries.shp")
shapes <- merge(shapes, dataset, by.x = 'ISO3', by.y = 'Country.Code')
```

## Plot data

We create a color pallete where teh countries with very high average population are depicted in Red.

```{r}
colors <- c("#8B0000","#FF0000", "#FFA07A")
bins <- c(100000, 20000, 5000, 0)
  
color_pal <- leaflet::colorBin(colors, 
                          domain = shapes$avgPopulationDensity, 
                          bins = bins)

leaflet(shapes) %>% 
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 1) %>%
  addPolygons(fillColor = ~color_pal(shapes$avgPopulationDensity),
              fillOpacity = 0.7,
              label = shapes$Country.Name,
              stroke = TRUE,
              color = 'black',
              weight = 1,
              opacity = 1)
```

We observe that countries in Asia like India, Bangladesh, China, Japan and others have high population densities than other countries.

